---
title: "Chapter 12 notes"
subtitle: "mutate() and case_when()"
author: "Kaelyn Lefferts"
date: today
toc: true
toc-depth: 2
number-sections: true
embed-resourses: true
format: html
---


# Quarto Notes
inline code: `{r} expression` 
```{r}
library(dplyr)
num_rows <- mtcars%>%
  filter(cyl==8)%>%
  nrow()
```
The filtered dataset has `{r}num_rows`

# Load Packages

```{r}
library(dslabs)
library(dplyr)
```

# Vectorized Conditionals
vectorized conditional functions evaluate conditions element-by-element and across a vector and return a new vector of the same length (different from subsetting which keeps only certain values)
```{r}
x <- 1:6
x>3 #logical vector using indexing
```
```{r}
x[x>3] #logical subsetting
```
```{r}
x[x>3] <- 99 #subassignment using a logical index
print(x)
```

## ifelse()
ifelse(test,yes,no), where test is the logical vector, yes is value to insert when TRUE, no is value to insert when test is FALSE
```{r}
x <- c(3,2,3,6,5)
length(x)
```
`x` is a vector with `{r} length(x)` values
```{r}
ifelse(test=x>3, yes="Large", no="Small")
length(ifelse(x>3, "Large", "Small"))
```

## case_when()
case_when(conditional1~ value1, condition2~ values2, .default=default_value)
~ means maps to, or yields
.default specifies the value to return when no conditions are TRUE, if omitted, case_when() will return NA for those positions
```{r}
case_when(c(3,2,3,6,5)>3~"Large", .default= "Small")
```
usually clearer once you have 3+ categories 

## More than Two Categories
```{r}
scores <- c(55,65,75,85,95)
```

### Nested ifelse()
```{r}
ifelse(scores<60, "Low",
       ifelse(scores<80, "Medium", "High"))
```
Works, but is confusing

### Multiple Rule case_when()
```{r}
case_when(scores<60~ "Low",
          scores<80 ~"Medium",
          TRUE~ "High")
```
Order of rules matters, case_when() uses the first condition that matches
```{r}
case_when(scores<80 ~ "Medium",
          scores<60 ~ "Low",
          TRUE ~ "High")
```
This does not work as intended, any score below 80 is labeled "Medium" before R ever checks the "Low" rule

# mutate()

## Column Copy
mutate(data, new_col= expression)
or name_data <- data %>% mutuate(new_col=expression)
take this df, and add a new column based on these existing columns
```{r}
names(mtcars)
```
```{r}
mtcars2 <- mtcars %>%
  mutate(new_column_copy=am) #identical to am column
```
```{r}
names(mtcars2)
```
Created a new column from an existing one

### mutate() with Arithmetic
```{r}
measurements <- data.frame(height_in=c(60,65,70))
print(measurements)
```
The `measurements` data frame has one column: `{r}colnames(measurements)`.
```{r}
measurements %>%
  mutate(height_cm=height_in*2.54)
```
Cool

### case_when() inside mutate()
```{r}
student_scores <- data.frame(scores=c(55,68,72,83,91))
student_scores <- student_scores %>%
  mutate(score_level=case_when(
    scores<60 ~ "Low",
    scores<80 ~ "Medium",
    TRUE ~ "High"
  ))
head(student_scores)
```
`student_scores` has the columns `{r} paste(names(student_scores), collapse = " and ")
Should render as the column names with an "and" in between

## Overwrite Column Values
```{r}
unique(mtcars$am)
```
```{r}
cars_overwrite <- mtcars %>%
  mutate(am=case_when(
    am==0 ~ "Automatic",
    TRUE ~ "Manual"
  ))
unique(cars_overwrite$am)
```
am column in cars_overwrite has been changes from 0/1 to "Automatic"/"Manual"
left side of = named the output column, right side of = created its values, if left side is a new name, R adds a column, if left side is an existing name, R overwrites that column
